Standalone Tactics using OpenTheory

### 2012-02-06 paper.tex
bit of work on sec 2
concrete, possibly running example, which might lead to a focus on
requirements of systems to be able to make use of standalone tactics

\subsection{\OpenTheory for Tactic Communication}
The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
The basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.

An article file represents a theory $\Gamma\rhd\Delta$.
As a special case, we can represent the proof of a single goal $g$ by taking $\Delta=\{\vdash g\}$.
More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem. \footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
(Additional inputs and outputs can always be communicated separately.)

As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
Such a task might be given to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic standalone. See <Sec 4??> for more realistic examples of tactics albeit in less detail.}

<more concrete examples? definition of tactics? viewing tactic as derived inference rule? intro to LCF kernel?>

>>>>

\subsection{\OpenTheory for Tactic Communication}
As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
The user of a proof assistant needs to prove this theorem, and decides to pass it off to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic stand alone. See \secref{performance} for more realistic examples of tactics (in less detail).}
The input for the tactic is the term $(\lambda{x}.\;x)\;\mathtt{T}$, and the output is a proof of the theorem $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
Standalone tactics use \OpenTheory article files for communicating both terms and theorems.
The proof assistant translates the term from its local internal format to \OpenTheory article format, and sends the resulting article over the web to the tactic.
The tactic returns another article, encoding a proof, which the proof assistant replays to create the desired theorem.

<Example illustrates requirements of the proof assistant: be able to read proofs from articles, be able to write terms as articles, and some form of communication with external program or website>

The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
The basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.

An article file represents a theory $\Gamma\rhd\Delta$.
As a special case, we can represent the proof of a single goal $g$ by taking $\Delta=\{\vdash g\}$.
More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem. \footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.
